Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-07-26 ms.tex

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby the action of self-gravity and radiative cooling alone produces galaxies with far too many stars. This problem has been addressed by employing various models of strong stellar feedback physics which are capable of generating self-regulating star formation in galaxies. \textbf{Energy injection from supernovae (SNe) has historically been used as the sole form of feedback. However, this is generally done heuristically as many simulations lack the ability to resolve the Sedov phase of individual SNe. But even with this strong feedback, recent work has argued for the need for pre-SN feedback, from stellar winds and/or stellar radiation, typically modelled simply as additional energy injection around newly formed stars (citations). Works that are capable of fulling resolving individual SN \citep{Hu2018,Smith2018} find that XXXXX Modelling these processes in detail, however, is challenging.

%Recently, it has become clear that the simplest models for stellar feedback, which account for supernovae (SNe) alone, are insufficient \citep{Smith2018} to fully regulate star formation and set galaxy properties consistent with observations (see \cite{NaabOstriker2017} for a recent review). Processes, such as stellar radiation, cosmic rays, or AGN, are often employed as additional sources of feedback. Modeling these processes in detail, however, is challenging. As a result, the relative importance of these additional physics is still uncertain.

Radiation from massive stars dominates the total feedback energy output of a stellar population \citep[e.g.][]{Abbott1982,Leitherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. %Lower mass stars (size range) primarily drive photoelectric heating (cite) and H$_2$ dissociation through radiation in the far ultraviolet and Lyman-Werner bands, while more massive stars (range) emit significant amounts of ionizing radiation. 
Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{WiseAbel2012,Kim2013a,Sales2014,Oshea2015,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs in four ways: 1) heating of diffuse gas and preventing the formation of cold, dense star formation regions, 2) destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

@@

In this work we use a detailed model for stellar feedback presented in \citep{Emerick2018} to study the role of radiation feedback in dwarf galaxy evolution. For the first time in a galaxy-scale simulation, we resolve the HII regions of single stars using an adaptive ray tracing radiative transfer method to follow the ionizing radaition from particles that represent individual stars. We focus on adressing two questions: 1) how important is radiation feedback in regulating star formation, and 2) are the long-range effects of radiation feedback important, or is the local destruction of dense gas the dominant effect. To investigate this we compare three simulations of the evolution of an isolated, low mass dwarf galaxy. Our fiducial model, containing full stellar radiation feedback, is compared against a run without ionziing radiation feedback, and a run with ionizing feedback limited to the local region around a given star. We dicuss our methods in Section~\ref{sec:methods} and present our results ion Section~\ref{sec:results}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{sfr}
\includegraphics[width=0.49\linewidth]{mass}
\caption{The star formation rate (left) and gas and stellar mass (right) of each of our three simulations over time. Time bins with no star formation are left empty. }
\label{fig:sfr_mass_evolution}
\end{figure*}

We compare the resulting star formation rate (left) and gas mass properties (right) of our three simulations in Figure~\ref{fig:sfr_mass_evolution}. There is a clear, significant contrast between the runs with and without ionizing radiation. Stellar ionizing radiation leads to a factor of $\sim$ 5 reduction in the resulting SFR, as compared to the noRT run. Since the shortrad and fiducial simulations are so similar over the first $\sim$~100~Myr, it is clear that stellar radiation acts to significantly reduce the local star formation efficiency around young, massive stars. Radiation from these stars quickly ionizes and dissipates surrounding dense gas that would otherwise have gone to fuel a significant amount of additional star formation. We are unable to follow the long-term evolution of the noRT galaxy due to computational constraints from the large number of stars formed. Although radiative transfer is computationally expensive, this run is substantially more expensive than the fiducial and shortrad runs \textbf{due to a lower typical timestep from more shock-heated SN gas and additional work computing the stellar wind and optically thin radiation effects from more star particles}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.98\linewidth]{density_temperature_ionization_z_profile}
\caption{Volume-averaged quantities above / below the disk of each galaxy. {\bf Left:} number density. {\bf Middle:} Temperature {\bf Right:} Hydrogen ionization fraction. These profiles correspond to the same point in time as Figure~\ref{fig:panel2}. \textbf{New figure. We can decide if we want to keep all panels, or stick with just 1 of the 3 quantities}}
\label{fig:vertical profiles}
\end{figure*}

\textbf{New text: Finally, we quantify these findings in Figure~\ref{fig:vertical profiles}, showing the volume-weighted average number density (left), temperature (middle) and hydrogen ionization fraction (right) as a function of distance above/below the disk of our galaxy at the same time as Figure~\ref{fig:panel2}. The gas outside the disk of our fiducial simulation is lower density, higher temperature, and has a higher ionization fraction at all points above the disk. The shortrad simulation, which lacks both the enhanced SNR of the noRT run and the long-range ionizing radiation effects, of the fiducial simulation, maintains a denser, cooler, more neutral shell of gas above the disk, preventing significant outflows.}

\textbf{New text: As shown in Figure~\ref{fig:metal_retention} the differences in ionization structure and its effect on galactic winds has direct consequences for the chemical evolution of our dwarf galaxy. The winds in our fiducial simulation carry nearly all of the metals produced out from the disk of our galaxy. This is the only run in agreement with observations of Local Group dwarf galaxies, both dSph's \citep{Kirby2011} and the gaseous, star forming Local Group dwarf galaxy Leo P \citep{McQuinn2015}, with metal ejection fractions of up to 95\%. This is in stark contrast to the $\sim$30\% retention fraction in our shortrad simulation. This would also influence the chemical enrichment of neighboring galaxies, given the significant differences in metal ejection past the virial radius between these two runs. Clearly the effects of feedback on observable chemical properties of galaxies is a key discriminator among models.}

>>>>

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby the action of self-gravity and radiative cooling alone produces galaxies with far too many stars. This problem has been addressed by employing various models of strong stellar feedback physics which are capable of generating self-regulating star formation in galaxies (see \cite{SomervilleDave2015} and \cite{NaabOstriker2017} for recent reviews). \textbf{Energy injection from supernovae (SNe) has historically been used as the sole form of feedback. However, this is generally done heuristically as many simulations lack the ability to resolve the Sedov phase of individual SNe. But even with this strong feedback, recent work has argued for the need for pre-SN feedback, from stellar winds and/or stellar radiation \citep[e.g.][]{Hu2016,Hopkins2018}, though this is typically modeled simply as additional energy injection around newly formed stars. The need for additional feedback is confirmed by simulations that are capable of fulling resolving individual SN \citep[e.g.]{Hu2017,Peters2017}. However, modelling these processes in detail is challenging, and their competing effects on galactic evolution is poorly constrained.}

Radiation from massive stars dominates the total feedback energy output of a stellar population \citep[e.g.][]{Abbott1982,Leitherer1999,Agertz2013}, surpassing the energy ejection of supernova ($\sim 10^{51}$~erg) by two orders of magnitude. If the radiation couples effectively to the interstellar medium (ISM), it can be a substantial source of additional feedback. Simulations including stellar radiation feedback followed through radiative transfer or radiation-hydrodynamics schemes have found it to be effective in regulating star formation and driving galactic winds \citep[e.g.][]{WiseAbel2012,Kim2013a,Sales2014,Oshea2015,Rosdahl2015,Ocvirk2015,Pawlik2015,Peters2017}. This occurs in four ways: 1) heating of diffuse gas and preventing the formation of cold, dense star formation regions, 2) destruction of cold, dense gas around recently formed stars, preventing further star formation, 3) momentum input by direct absorption of UV radiation by gas and (in some cases) dust through re-emission and scattering in the infrared, and 4) lowering the typical ISM densities in which SNe occur and greatly increasing their effectiveness. % the relevance / importance of RP from dust in the IR is beyond the scope of this work...

@@

In this work we use a detailed model for stellar feedback presented in \cite{Emerick2018} to study the role of radiation feedback in dwarf galaxy evolution. For the first time in a galaxy-scale simulation, we resolve the HII regions of single stars using an adaptive ray tracing radiative transfer method to follow the ionizing radiation from particles that represent individual stars. We focus on addressing two questions: 1) how important is radiation feedback in regulating star formation, and 2) are the long-range effects of radiation feedback important, or is the local destruction of dense gas the dominant effect. To investigate this we compare three simulations of the evolution of an isolated, low mass dwarf galaxy. Our fiducial model, containing full stellar radiation feedback, is compared against a run without ionizing radiation feedback, and a run with ionizing feedback limited to the local region around a given star. We discuss our methods in Section~\ref{sec:methods} and present our results ion Section~\ref{sec:results}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{sfr}
\includegraphics[width=0.49\linewidth]{mass}
\caption{The star formation rate (left) and gas and stellar mass (right) of each of our three simulations over time. Time bins with no star formation are left empty.}
\label{fig:sfr_mass_evolution}
\end{figure*}

We compare the resulting star formation rate (left) and gas mass properties (right) of our three simulations in Figure~\ref{fig:sfr_mass_evolution}. There is a clear, significant contrast between the runs with and without ionizing radiation. Stellar ionizing radiation leads to a factor of $\sim$ 5 reduction in the resulting SFR, as compared to the noRT run. Since the shortrad and fiducial simulations are so similar over the first $\sim$~100~Myr, it is clear that stellar radiation acts to significantly reduce the local star formation efficiency around young, massive stars. Radiation from these stars quickly ionizes and dissipates surrounding dense gas that would otherwise have gone to fuel a significant amount of additional star formation. We are unable to follow the long-term evolution of the noRT galaxy due to computational constraints from the large number of stars formed. Although radiative transfer itself is computationally expensive, this run is substantially more costly \textbf{due to a lower typical timestep and increased cost in computing the optically thin radiation effects (photoelectric effect and LW dissociation).}

@@

%\begin{figure*}
%\centering
%\includegraphics[width=0.98\linewidth]{density_temperature_ionization_z_profile}
%\caption{Volume-averaged quantities above / below the disk of each galaxy. {\bf Left:} number density. {\bf Middle:} Temperature {\bf Right:} Hydrogen ionization fraction. These profiles correspond to the same point in time as Figure~\ref{fig:panel2}. \textbf{New figure. We can decide if we want to keep all panels, or stick with just 1 of the 3 quantities}}
%\label{fig:vertical profiles}
%\end{figure*}

%\textbf{Finally, we quantify these findings in Figure~\ref{fig:vertical profiles}, showing the volume-weighted average number density (left), temperature (middle) and hydrogen ionization fraction (right) as a function of distance above/below the disk of our galaxy at the same time as Figure~\ref{fig:panel2}. The gas outside the disk of our fiducial simulation is lower density, higher temperature, and has a higher ionization fraction at all points above the disk. The shortrad simulation, which lacks both the enhanced SNR of the noRT run and the long-range ionizing radiation effects, of the fiducial simulation, maintains a denser, cooler, more neutral shell of gas above the disk, preventing significant outflows.}

\textbf{As shown in Figure~\ref{fig:metal_retention} the differences in ionization structure and its effect on galactic winds has direct consequences for the chemical evolution of our dwarf galaxy. The winds in our fiducial simulation carry nearly all of the metals produced out from the disk of our galaxy. This is the only run in agreement with observations of Local Group dwarf galaxies, both dSph's \citep{Kirby2011} and the gaseous, star forming Local Group dwarf galaxy Leo P \citep{McQuinn2015}, with metal ejection fractions of up to 95\%. This is in stark contrast to the $\sim$30\% retention fraction in our shortrad simulation. This would also influence the chemical enrichment of neighboring galaxies, given the significant differences in metal ejection past the virial radius between these two runs. Clearly the effects of feedback on observable chemical properties of galaxies is a key discriminator among models.}
